### 2014-02-11 classicthesis/Chapters/Background.tex
alle tekst die niet van mij was is nu echt weg :)

%************************************************
\chapter{Introduction to Learning-to-Rank}
\chapter{Benchmark Datasets}
\chapter{Pointwise Approach}
\chapter{Pairwise Approach}
\chapter{Listwise Approach}
\chapter{Summary}