### 2014-02-11 classicthesis/FrontBackmatter/Acknowledgments.tex
references fixed :)

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape    
    We have seen that computer programming is an art, \\ 
    because it applies accumulated knowledge to the world, \\ 
    because it requires skill and ingenuity, and especially \\
    because it produces objects of beauty.} \\ \medskip
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
Put your acknowledgments here.

Many thanks to everybody who already sent me a postcard!

Regarding the typography and other help, many thanks go to Marco 
Kuhlmann, Philipp Lehman, Lothar Schlesier, Jim Young, Lorenzo 
Pantieri and Enrico Gregorio\footnote{Members of GuIT (Gruppo 
Italiano Utilizzatori di \TeX\ e \LaTeX )}, J\"org Sommer, 
Joachim K\"ostler, Daniel Gottschlag, Denis Aydin, Paride 
Legovini, Steffen Prochnow, Nicolas Repp, Hinrich Harms, 
 Roland Winkler, J\"org Weber, 
 and the whole \LaTeX-community for support, ideas and 
 some great software.

\bigskip

\noindent\emph{Regarding \mLyX}: The \mLyX\ port was intially done by 
\emph{Nicholas Mariette} in March 2009 and continued by 
\emph{Ivo Pletikosi\'c} in 2011. Thank you very much for your 
work and the contributions to the original style.

\endgroup