### 2014-02-12 classicthesis/FrontBackmatter/Acknowledgments.tex
acknowledgement geschreven en dedication aangepast

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape    
    We have seen that computer programming is an art, \\ 
    because it applies accumulated knowledge to the world, \\ 
    because it requires skill and ingenuity, and especially \\
    because it produces objects of beauty.} \\ \medskip
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their valuable insights and great supervision.

In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees who were so kind to provide me with valuable insights.

\endgroup