### 2014-02-12 classicthesis/FrontBackmatter/Acknowledgments.tex
mooie acknowledgement quote :)

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape    
   God gave you a gift of 86 400 seconds today. \\
   Have you used one to say thank you?.} \\ \medskip
   - William Arthur Ward
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their great supervision throughout the project.

In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine or elsewhere. In particular I owe many thanks to Michel Barneveld for giving me insight in potential applications of Learning-to-Rank within Avanade.

\endgroup