### 2014-02-12 classicthesis/BenchmarkResults.tex
structuurswijziging + parttext

\chapter{Yahoo! Learning to Rank Challenge}
\section{Benchmark characteristics}
\section{Results}
\chapter{Yandex Internet Mathematics competition}
\section{Benchmark characteristics}
\section{Results}
\chapter{LETOR}
\section{Benchmark characteristics}
\section{Results}
\chapter{MSLR-WEB30k}
\section{Benchmark characteristics}
\section{Results}