### 2014-05-20 azure_grant_proposal/proposal.tex
opmaak

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012
\documentclass{sig-alternate}
\usepackage{array}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{booktabs}
\begin{document}

%
% --- Author Metadata here ---
\conferenceinfo{}{}
\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Research Proposal: A Scalability Analysis of Learning to Rank Algorithms using Hadoop}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Niek Tax\\
       \affaddr{University of Twente}\\
       \affaddr{P.O. Box 217}\\
       \affaddr{7500 AE Enschede}\\
       \affaddr{The Netherlands}
       \affaddr{d.hiemstra@utwente.nl}
}

\maketitle

\begin{abstract}
Learning to rank is an increasingly important scientific field that comprises the use of machine learning for the ranking task. This research proposal proposes a two-stage research project in the learning to rank field. In the first research stage we will conduct a comparison of learning to rank methods over multiple benchmark collections. This is relevant, as a lack of a \emph{de facto} standard benchmark collection for evaluation of new learning to rank methods hinders comparison. The second stage analyses the potential speed-up gain of parallel execution of the most accurate learning to rank methods identified in the first research step through execution on Hadoop MapReduce. Research on scalability of learning to rank methods is widely seen as relevant and often overlooked \cite{Chapelle2011b, Liu2007}.
\end{abstract}
% A category with the (minimum) three required fields
\category{H.3.3}{Information Systems}{Information Storage and Retrieval}[Information Search and Retrieval]
%
\category{D.1.3}{Concurrent Programming}{Distributed Programming}

\keywords{Learning to Rank, MapReduce}
\section{Introduction}
Ranking is a core problem in the field of information retrieval. The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query. Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web. One can find useful applications of ranking in many application domains outside web search as well. For example, it plays a vital role in amongst others: automatic document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}. In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\

Research in the field of ranking models has for a long time been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}. Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957. The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models. Learning to rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\

In recent years several learning to rank benchmark datasets have been proposed that enable comparison of the performance of different Learning to rank methods. Well-known benchmark datasets include the \emph{Yahoo! Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was published by Microsoft Research. One of the concluding observations of the \emph{Yahoo! Learning to Rank Challenge} was that almost all work in the Learning to rank field focuses on ranking accuracy, while efficiency and scalability of learning to rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}. Liu \cite{Liu2007} confirms the observation that efficiency and scalability of learning to rank methods has so far been an overlooked research area in his influential book on learning to rank.\\

Some research has been done in the area of parallel or distributed machine learning \cite{Chu2007,Chang2007}. However, almost none of these studies include the learning to rank sub-field of machine learning. The field of efficient learning to rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007. Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient learning to rank through the use of parallel programming paradigms.\\

MapReduce \cite{Dean2004} is a parallel programming framework that is inspired by the \emph{Map} and \emph{Reduce} functions commonly used in functional programming. Since Google developed the MapReduce parallel programming framework back in 2004 it has grown to be the industry standard model for parallel programming. Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most learning to rank algorithms are, are not amenable to the MapReduce framework. Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\

The appearance of benchmark datasets gave insight in the performance of different learning to rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets. Up to now it remains unknown whether popular existing learning to rank methods scale well when they are used in a parallel manner using the MapReduce framework. This thesis aims to be an exploratory start in this little researched area of parallel learning to rank.\\

\section{Research Goals}
\label{sec:goals}
Our research contains two research stages. A first stage is concerned with identifying the best performing learning to rank methods in terms of accuracy. Several benchmark collections for model evaluation exist in the learning to rank field. Well-known benchmark collections include the \emph{Yahoo! Learning to Rank Challenge} collection \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}. There exists no agreement among authors in the learning to rank field on which benchmark collection(s) to use to evaluate a new model. Comparing ranking accuracy of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models. Therefore, our first research question concerns the identification of the most accurate learning to rank methods when compared on their performance on several benchmark collections.
\begin{description}
\item[RQ1] What are the best performing learning to rank algorithms in terms of accuracy on when compared over multiple benchmark collections?
\end{description}
The second stage of our research is concerned with exploration of the speed-up in execution time of learning to rank algorithms through parallelisation using the Hadoop MapReduce framework. This second stages focuses on those learning to rank algorithms that showed to be most accurate in our cross-benchmark comparison of learning to rank methods.
\begin{description}
\item[RQ2] What is the speed-up of those learning to rank algorithms when executed using the MapReduce framework?
\end{description}
\noindent Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\

$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$

\section{Methodology}
The following subsections will describe the research methodology per stage of our two-stage research.
\subsection{Cross-benchmark learning to rank comparison}
We collect evaluation results on the datasets of benchmark collections through a structured literature search. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
\begin{itemize}\itemsep0em
\item LETOR 2.0
\item LETOR 3.0
\item LETOR 4.0
\item Yahoo! Learning to Rank Challenge
\item Yandex Internet Mathematics 2009 contest
\item MSLR-web10/30k
\item WCL2R
\item AOL
\end{itemize}

For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\

LETOR 1.0 and 3.0, Yahoo! Learning to Rank Challenge, WCL2R and AOL have accompanying papers which were published together with these benchmark collections. Authors publishing evaluation results on these benchmark collections are requested to cite these papers. Therefore, we collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search. Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search. Google Scholar will be used to perform the forward reference search.

\begin{table}[!h]
\begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
Benchmark & Paper & \# of forward references \\
\midrule
LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
AOL dataset & Pass et al. \cite{Pass2006} & 339\\
WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
\bottomrule
\end{tabular}
\caption{Forward references of learning to rank benchmark papers}
\label{tbl:ltr_benchmark_forref}
\end{table}

The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections are not accompanied by a paper. To collect evaluation results for learning to rank methods on these benchmarks, a Google Scholar search is performed on the name of the benchmark. Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.

\begin{table}[!h]
\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
Search Query & Google scholar search results \\
\midrule
"LETOR 4.0" & 75 \\
"MSLR-web10k" & 16 \\
"MSLR-web30k" & 15 \\
"Yandex Internet Mathematics" & 3 \\
\bottomrule
\end{tabular}
\caption{Google scholar search results for learning to rank benchmarks}
\label{tbl:ltr_benchmark_searches}
\end{table}

\subsection{Analysis of scalability on MapReduce}
The learning to rank methods identified in the first research stage will be implemented in Pig to run them in parallel on a MapReduce platform. The Microsoft HDInsight cloud-based MapReduce implementation will be used to execute to run the experiments on. Microsoft HDInsight has the benefit of being able to scale to large numbers of MapReduce nodes with guaranteed availability.

\bibliographystyle{abbrv}
\bibliography{Bibliography}  % sigproc.bib is the name of the
\end{document}