### 2014-09-22 report/FrontBackmatter/Preface.tex
herstructurering frontbackmatter

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Preface}{preface}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Preface}
Dear reader,\\

Thank you for taking interest in this thesis, which I have written as final project as part of the Master programme in Computer Science at the University of Twente. This research was conducted at Avanade Netherlands B.V. under the primary supervision of Sander Boackting Msc. I would like to use this page to express my gratitude to everyone who supported my throughout this project in any way.\\

Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their great supervision throughout the project.\\

In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine, during the Friday afternoon drink, or elsewhere. In particular I owe many thanks to Michel Barneveld for giving me insight in potential applications of Learning-to-Rank within Avanade.\\

-- Niek Tax

\endgroup