### 2014-09-29 report/FrontBackmatter/Titlepage.tex

%*******************************************************
% Titlepage
%*******************************************************
\begin{titlepage}
\title{Master's thesis Computer Science\\ (Chair: Databases, Track: Information System Engineering)}
\course{Scaling Learning to Rank to Big Data}
\subtitle{A study concerning parallelisation of Learning to Rank algorithms using the MapReduce computing model}
\university{University of Twente}
\faculty{Faculty of Electrical Engineering, Mathematics and Computer Science (EEMCS)}
\author{Niek Tax (s0166197)}

\committee{
Dr. ir. Djoerd Hiemstra (University of Twente / Databases)\\
Dr. ir. Dolf Trieschnigg (University of Twente / Human Media Interaction)\\
Sander Bockting, Msc (Avanade Netherlands B.V.)\\
}	
\date{\today}
\num{}

\supervisors{
Dr. ir. Djoerd Hiemstra (University of Twente / Databases)\\
Sander Bockting, Msc (Avanade Netherlands B.V.)
}    
\end{titlepage}
\maketitle
\clearpage