### 2014-09-29 report/FrontBackmatter/Preface.tex
Dolf er ook in

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Preface}{preface}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Preface}
Dear reader,\\

Thank you for taking interest in this thesis, which I have written as final project as part of the Master programme in Computer Science at the University of Twente. This research was conducted at Avanade Netherlands B.V. under the primary supervision of Sander Bockting Msc of Avanade and Dr. ir. Djoerd Hiemstra from the University of Twente. I would like to use this page to express my gratitude to everyone who supported me throughout this project in any way.\\

Many thanks go to Dr. ir. Djoerd Hiemstra from the University of Twente and to Sander Bockting Msc of Avanade Netherlands B.V. for their great supervision throughout the project. Even though we kept face-to-face meetings to a very minimum, you both provided me with very insightful and valuable feedback either in those meetings or per e-mail. Also, I would like to thank Dr. ir. Dolf Trieschnigg, together with Djoerd and Sander a member of the assessment committee of this graduation project, for still being available for the role as second assessor of this work, while having joined the process rather late.\\

In addition I would like to thank all fellow graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine, during the Friday afternoon drink, or elsewhere. In particular I would like to mention fellow graduate interns Fayaz Kallan, Casper Veldhuijzen, Peter Mein, and Jurjen Nienhuis for the very good time that we had together at the office as well as during the numerous drinks and dinners that we had together outside office hours.\\

I finish this section by thanking everyone that helped improving the quality of my work by providing me with valuable feedback. I would like to thank my good friends and former fellow boards members at study association I.C.T.S.V. Inter-\emph{Actief} Rick van Galen and Jurri\"{e}n Wagenaar, who provided me with feedback in the early stages of the process. In particular I would like to thank Jurjen Nienhuis (yes, again), for the numerous mutual feedback sessions that we held, which most certainly helped raising the quality of this thesis to a higher level.\\
\bigskip

-- Niek Tax

\endgroup