### 2014-09-30 report/FrontBackmatter/Titlepage.tex
generalisatie van titlepage

%*******************************************************
% Titlepage
%*******************************************************
\begin{titlepage}
\title{Master's thesis Computer Science\\ (Chair: \myDepartment, Track: \myTrack)}
\course{\myTitle}
\subtitle{\mySubtitle}
\university{\myUni}
\faculty{\myFaculty}
\author{\myName (\myStudNum)}

\committee{
\myProf (\myProfAffil)\\
\myOtherProf (\myOtherProfAffil)\\
\mySupervisor (\mySupervisorAffil)\\
}	
\date{\today}
\num{}

\supervisors{
\myProf (\myProfAffil)\\
\mySupervisor (\mySupervisorAffil)
}    
\end{titlepage}
\maketitle
\clearpage